# Chapter 1. Data Visualization with ggplot2

```{r,warning=FALSE,message=FALSE}
library("tidyverse")

## Engine Size vs Car Fule Efficiency

ggplot(data = mpg) + 
    geom_point(mapping = aes(x = displ, y = hwy))

## The plot shows a negative relationship between engine size (displ) and fuel efficiency (hwy)

```

With ggplot2, you begin a plot with the function ggplot(). The first argument of ggplot() is the dataset to use in the graph. The function geom_point() adds a layer of points to your plot, which creates a scatterplot.

### A Graphing Template

ggplot(data = <DATA>) +
  <GEOM_FUNCTION>(mapping = aes(<MAPPINGS>))
  
### Aesthetic Mappings

```{r,warning=FALSE,message=FALSE}

## showing different color for each class
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, color = class))

## We can use size as well, but using size for a non ordered variable is not advised

ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, size = class))

## we can also use alpha for transperency and shape for different shapes

ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, alpha = class))

ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, shape = class)) ## only 6 shapes

```

#### We can also set aesthetics manually.

```{r,warning=FALSE,message=FALSE}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy), color = "blue")

```

* The name of a color as a character string
* The size of a point in mm
* The shape of a point as a number

### Facets

Another way, particularly useful for categorical variables, is to split your plot into facets, subplots that each display one subset of the data.

```{r,warning=FALSE,message=FALSE}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_wrap(~ class, nrow = 2)

```

To facet your plot on the combination of two variables, add facet_grid() to your plot call. The first argument of facet_grid() is also a formula. This time the formula should contain two variable names separated by a ~

```{r,warning=FALSE,message=FALSE}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_grid(drv ~ cyl)

```

### Geometric Objects

```{r,warning=FALSE,message=FALSE}
## Just a scatter plot
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy))

## Smooth lines.
ggplot(data = mpg) +
  geom_smooth(mapping = aes(x = displ, y = hwy))

# Here geom_smooth() separates the cars into three lines based on their drv value, which describes a car’s drivetrain.
ggplot(data = mpg) +
  geom_smooth(mapping = aes(x = displ, y = hwy, linetype = drv))

## Overlaying points and smooth line over each other
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, color = drv)) +
    geom_smooth(mapping = aes(x = displ, y = hwy, linetype = drv))

```


There's duplication in previous set of code. We can set aesthetics at global level

```{r,warning=FALSE,message=FALSE}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = drv)) +
  geom_point() +
    geom_smooth()

```

If you place mappings in a geom function, ggplot2 will treat them as local mappings for the layer. It will use these mappings to extend or overwrite the global mappings for that layer only. This makes it possible to display different aesthetics in different layers

```{r,warning=FALSE,message=FALSE}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) +
  geom_point(mapping = aes(color = class)) +
  geom_smooth()

```

You can use the same idea to specify different data for each layer. Here, our smooth line displays just a subset of the mpg dataset, the subcompact cars. The local data argument in geom_smooth() overrides the global data argument in ggplot() for that layer only

```{r,warning=FALSE,message=FALSE}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) +
  geom_point(mapping = aes(color = class)) +
  geom_smooth(
    data = filter(mpg, class == "subcompact"),se = FALSE
  )
```

### Statistical Transformations

* Bar charts, histograms, and frequency polygons bin your data and then plot bin counts, the number of points that fall in each bin.

* Smoothers fit a model to your data and then plot predictions from the model.

* Boxplots compute a robust summary of the distribution and display a specially formatted box

```{r,warning=FALSE,message=FALSE}
ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = cut))
```

You can generally use geoms and stats interchangeably. For example, you can re-create the previous plot using stat_count() instead of geom_bar()

```{r,warning=FALSE,message=FALSE}
ggplot(data = diamonds) +
  stat_count(mapping = aes(x = cut))
```


I change the stat of geom_bar() from count (the default) to identity, this lets me map the height of the bars to the raw values of a y variable

```{r,warning=FALSE,message=FALSE}
demo <- tribble(
  ~a,      ~b,
  "bar_1", 20,
  "bar_2", 30,
  "bar_3", 40
)

ggplot(data = demo) +
  geom_bar(
    mapping = aes(x = a, y = b), stat = "identity"
  )

```

You might want to override the default mapping from transformed variables to aesthetics. For example, you might want to display a bar chart of proportion, rather than count

```{r,warning=FALSE,message=FALSE}
ggplot(data = diamonds) +
  geom_bar(
    mapping = aes(x = cut, y = ..prop.., group = 1)
  )

```

You might want to draw greater attention to the statistical transformation in your code. For example, you might use stat_summary(), which summarizes the y values for each unique x value, to draw attention to the summary that you’re computing:

```{r,warning=FALSE,message=FALSE}
ggplot(data = diamonds) +
  stat_summary(
    mapping = aes(x = cut, y = depth),
    fun.ymin = min,
    fun.ymax = max,
    fun.y = median
  )

```

ggplot2 provides over 20 stats for you to use. Each stat is a function, so you can get help in the usual way, e.g., ?stat_bin. To see a complete list of stats, try the ggplot2 cheatsheet.

### Position Adjustments

You can color a bar chart using either the color aesthetic, or more usefully, fill

```{r,warning=FALSE,message=FALSE}
## Color is used for borders
ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = cut, color = cut))

## fill, fills the color
ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = cut, fill = cut))

## When we use fill to other variable

ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = cut, fill = clarity))


ggplot(
  data = diamonds,
  mapping = aes(x = cut, fill = clarity)
) +
  geom_bar(alpha = 1/5, position = "identity")

ggplot(
  data = diamonds,
  mapping = aes(x = cut, color = clarity)
) +
  geom_bar(fill = NA, position = "identity")
```

* Position = 'identity': will place each object exactly where it falls in the context of the graph. This is not very useful for bars, because it overlaps them.

```{r,warning=FALSE,message=FALSE}

ggplot(
  data = diamonds,
  mapping = aes(x = cut, fill = clarity)
) +
  geom_bar(alpha = 1/5, position = "identity")

ggplot(
  data = diamonds,
  mapping = aes(x = cut, color = clarity)
) +
  geom_bar(fill = NA, position = "identity")
```

* position = "fill" works like stacking, but makes each set of stacked bars the same height. This makes it easier to compare proportions across groups:

```{r,warning=FALSE,message=FALSE}

ggplot(data = diamonds) +
  geom_bar(
    mapping = aes(x = cut, fill = clarity),
    position = "fill"
  )
```

* position = "dodge" places overlapping objects directly beside one another. This makes it easier to compare individual values:

```{r,warning=FALSE,message=FALSE}

ggplot(data = diamonds) +
  geom_bar(
    mapping = aes(x = cut, fill = clarity),
    position = "dodge"
  )
```


### Co-ordinate sytems

* coord_flip() switches the x- and y-axes. 

```{r,warning=FALSE,message=FALSE}

ggplot(data = mpg, mapping = aes(x = class, y = hwy)) +  geom_boxplot()

ggplot(data = mpg, mapping = aes(x = class, y = hwy)) +  geom_boxplot() +  coord_flip()
```

End of Chapter 1 -- 