## Chapter 2: Workflow Basics

* R Can be used as calculator
* We use "<-" as assignment operator. It is referred as "gets"

## Chapter 3: Data Transformation with dplyr

* filter()


```{r,message=FALSE, warning=FALSE}
library(nycflights13)
library(tidyverse)

flights %>% filter(month == 1, day == 1)

## Logical
flights %>% filter(month == 11| month == 12)

## in clause
flights %>% filter(month %in% c(11,12))

```

* arrange()

```{r,message=FALSE, warning=FALSE}

flights %>% arrange(year, month, day)

```

* select()

```{r,message=FALSE, warning=FALSE}

flights %>% select(year, month, day)

```

* rename()

```{r,message=FALSE, warning=FALSE}

flights %>% rename(tail_num = tailnum)

```

* mutate()

```{r,message=FALSE, warning=FALSE}

flights %>% mutate(hours = air_time/60)

##transmutate will keep only new variables

flights %>% transmute(hours = air_time/60)

```

* offset

lead() and lag() allow you to refer to leading or lagging values. This allows you to compute running differences (e.g., x - lag(x)) or find when values change (x != lag(x)). They are most useful in conjunction with group_by()

```{r,message=FALSE, warning=FALSE}

(x <- 1:10)
lag(x)
lead(x)

```

* Cumulative and rolling aggregates

```{r,message=FALSE, warning=FALSE}

cumsum(x)
cummean(x)
cummin(x)
cummax(x)

```

* Ranking

```{r,message=FALSE, warning=FALSE}

y <- c(1, 2, 2, NA, 3, 4)
min_rank(y)
min_rank(desc(y))

dense_rank(y)
```

* Summarize()


